### 2015-07-07 narrative/data-description.tex

Users will also need to describe information about the data in order to choose
the optimal data-refactoring routine. Currently, we understand that
users can tell us that the data represents: 1)~phase-space (e.g., the particles
in our example), 2)~spatial-changes (e.g., the mesh), 3)~space-time (e.g.,
the fields on the mesh). We need this information to select
the best possible data
refactoring methods. Although users can insert a user-defined data-refactoring
method, system-level methods can also use this information to make better decisions.
We will investigate new techniques and new semantics capable of using
this information to best prioritize and re-organize data.

Finally we will investigate the use of additional information
on top of the data model.
This information will be used to indicate data associations that exceed
traditional data models. For example, the fields and mesh in our example have
an inherent relationship. A common data model will describe all of the
variables in the mesh (the coordinates, the units, and the types of elements if
this were a finite-element mesh). We will explore new mechanisms that will
group data together making it explicit that M1, F1, and G1 need to be organized
together.  In other words, it will not be helpful if data from M1 and M2 were
organized together along with only F1. When the users wants to visualize data
from the field, F1, they will need only M1 and M2 would be additional
information. Likewise, if the user wants to see F1 and F2, then they will need
the mesh from M1 and M2. This requires that the time to access all of this
data is approximately the same requiring similar access times for all of the
data.  We will investigate what choices the SSIO system must make in order to
keep these pieces ``together'', from a performance perspective. Can the storage
system move everything to a lower layer when all of it doesn't fit or is there
other, less tightly related data can be evicted to preserve the collective
retrieval performance?  We need to investigate the potential problems when this
occurs and how the users can specify additional information to ensure they get
the best possible data from the SSIO layer.

>>>>

Users will also need to describe information about the data in order to choose
the optimal data-refactoring routine. Currently, we understand that
users can tell us that the data represents: 1)~phase-space (e.g., the particles
in our example), 2)~spatial-changes (e.g., the mesh), and 3)~space-time (e.g.,
the fields on the mesh). We need this information to select
the best possible data
refactoring methods. Although users can insert a user-defined data-refactoring
method, system-level methods can also use this information to make better decisions.
We will investigate new techniques and new semantics capable of using
this information to best prioritize and re-organize data.

Finally we will investigate the use of additional information
on top of the data model.
This information will be used to indicate data associations that exceed
traditional data models. For example, the fields and mesh in our example have
an inherent relationship. A common data model will describe all of the
variables in the mesh (the coordinates, the units, and the types of elements if
this were a finite-element mesh). We will explore new mechanisms that will
group data together making it explicit that M1, F1, and G1 need to be organized
together.  In other words, it will not be helpful if data from M1 and M2 were
organized together along with only F1. When the users wants to visualize data
from the field, F1, they will need only M1, and M2 would be additional
information. Likewise, if the user wants to see F1 and F2, then they will need
the mesh from M1 and M2. This requires that the time to access all of this
data is approximately the same.
We will investigate what choices the SSIO system must make in order to
keep these pieces ``together'', from a performance perspective. Can the storage
system move everything to a lower layer when all of it doesn't fit or is there
other, less tightly related data that can be evicted to preserve the collective
retrieval performance?  We need to investigate the potential problems when this
occurs and how the users can specify additional information to ensure they get
the best possible data from the SSIO layer.

### 2015-07-07 narrative/deliverables.tex

\item Assemble and evaluate general techniques for refactoring of data
that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
\item Explore the use of application hints and utility of data to guide the initial placement of data. 
(Rutgers/ORNL)
\item Investigate the trade-off between ``filing'' and ``piling'' data. (UCSC/Sandia)
\item Demonstrate a metadata service capable of serving both POSIX clients and our clients,
but without consideration for authentication, authorization, or data migration. (Sandia/UCSC)
\end{itemize}

\textbf{Year 2}
\begin{itemize}
\item Develop a new type of querying system that allows an application to ask the storage system
about completion timing information. (ORNL)
\item Research autonomic data management strategies that
can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime.
\item Demonstrate a time bounded search approach for finding data within the storage system
to identify data and the current location. This will update the cached information in the metadata
service to short-circuit future requests for the data presuming that it does not migrate again soon. 
\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions.
(Sandia/UCSC)
\end{itemize}

\textbf{Year 3}
\begin{itemize}
\item Work with DOE applicationexsiting ADIOS-enabled applications, framework using many of the applications to which we have access, including XGC1, GTC, and SPECFM3D
%\item storage APIs to manipulate storing and retrieving data to specific tiers
%\item metadata management for data stored in multiple tiers using different compression
%\item metadata management for deeper data knowledge
%\item middleware hooks to manage new storage APIs
%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}

>>>>

\item Assemble and evaluate general techniques for refactoring of data
that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
\item Explore the use of application hints and utility of data to guide the initial placement of data. 
(Rutgers/ORNL)
\item Investigate the trade-off between ``filing'' and ``piling'' data. 
Demonstrate a time bounded search approach for finding data within the storage system
to identify data and the current location. (UCSC/Sandia)
\item Demonstrate a metadata service capable of serving both POSIX clients and our clients. (Sandia/UCSC)
\end{itemize}

\textbf{Year 2}
\begin{itemize}
\item Develop a new type of querying system that allows an application to ask the storage system
about completion timing information. (ORNL)
\item Evaluate the effectiveness and overhead to re-organize data and use the multi-layer approach 
to staging. (Brown/ORNL)
\item Research autonomic data management strategies that
can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime. 
Research runtime tracking and estimation. (Rutgers/ORNL)
\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions. (Sandia)
(Sandia/UCSC)
\end{itemize}

\textbf{Year 3}
\begin{itemize}
\item Work with DOE applications, particularly XGC1, GTC, and SPECFM3D, to assess 
the new I/O interface including both APIs and hints. (ORNL)
\item Research techniques that can enhance migration without making the eventual 
lookup of data unbounded, and manage each tier in a scalable fashion. (ORNL)
\item Demonstrate incorporating A\&A and show scalability under both application load as
well as storage system pressures showing that we can tolerate both loads and maintain quality of
service. (Sandia)
%\item storage APIs to manipulate storing and retrieving data to specific tiers
%\item metadata management for data stored in multiple tiers using different compression
%\item metadata management for deeper data knowledge
%\item middleware hooks to manage new storage APIs
%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}
